Field trial of programmable L3 VPN service deployment using SDN-Based Multi-domain Service Provisioning over IP/Optical networks

### 2021-05-14 main.tex
update

\author{
\IEEEauthorblockN{Samier Barguil\IEEEauthorrefmark{1}, Victor Lopez\IEEEauthorrefmark{2}, Cristyan Manta-Caro\IEEEauthorrefmark{4}, Arturo Mayoral Lopez De Lerma\IEEEauthorrefmark{2},\\
Oscar Gonzalez De Dios\IEEEauthorrefmark{2}, Edward Echeverry\IEEEauthorrefmark{3}, Juan Pedro Fernandez-Palacios\IEEEauthorrefmark{2}, Janne Karvonen\IEEEauthorrefmark{5},\\
Jutta Kemppainen\IEEEauthorrefmark{5}, Natalia Maya\IEEEauthorrefmark{5}, and Ricard Vilalta\IEEEauthorrefmark{6}}
\IEEEauthorblockA{
\IEEEauthorrefmark{1}Universidad Autonoma de Madrid, Madrid, Spain\\
\IEEEauthorrefmark{2}Telefonica I+D, Ronda de la Comunicacion, Madrid, Spain \\
\IEEEauthorrefmark{3}Telefonica Movistar, Transversal 60 No 114ª -55. Bogotá, Colombia\\
\IEEEauthorrefmark{4}Wipro Technologies Ltd., Doddakannelli, Sarjapur Road
Bengaluru - 560 035, India\\
\IEEEauthorrefmark{5}Infinera Corporation, 140 Caspian Court, Sunnyvale, CA 94089, USA\\
\IEEEauthorrefmark{6}Centre Tecnol\`{o}gic de Telecomunicacions de Catalunya (CTTC/CERCA), Castelldefels, Spain}

}

@@

\begin{abstract}
%Software-Defined Networking (SDN) is a powerful paradigm already transforming the everyday operations in Telecommunications Networks. SDN came with the idea of decoupling forwarding and control plane in the switches via OpenFlow, but it has evolved to cope with the needs of production networks. 
Network operators can not change their footprint to upgrade the entire Network to support SDN from scratch. \replaced{Hence,}{This why} network operators adapt\added{ed} the original SDN concept into a hybrid SDN approach to have a pragmatic, evolutionary, and economically viable solution. This paper tests an SDN architecture based on a hierarchical structure of a Software-Defined Transport Networking (SDTN) controller, which deals with the end-to-end service provision and topology collection on top of a set of SDN controllers for each technological domain. The authors demonstrate L3/L2 VPN service deployment using an underlay multi-domain IP/Optical networks in a field trial. The approach allows a service provider to migrate brownfield scenarios into SDN domains using programmatic interfaces.

%network islands into SDN domains using programmatic interfaces.

%Service provisioning is a key process in the value chain for supplying next-generation services to customers of all sizes and characteristics. Commonly, during years service provisioning was executed manually, then supported by service activators tuned tools for the vendor- specific combination of network elements. With the advent boom of SDN, service delivery operations can be performed in an agnostic-vendor fashion using standard data models and protocols. However, new challenges still persist such as orchestrate multiple layers required for covering long-haul, medium and short distances. Multidomain networking between IP/MPLS-based layers and underlying WDM multi-layer technologies require further coordination and orchestration. This presents two use cases to enable multidomain service provisioning and the corresponding topology visualization. Moreover, this work demonstrates the capabilities of a SDTN controller in a field trial environment our approach for Hybrid SDN architectures.
\end{abstract}

%\begin{IEEEkeywords}
%Security SLA, NFV, orchestration, V2X, C-ITS, CAM, DENM, ACCA
%\end{IEEEkeywords}

%\vspace{-0.2in}

@@

%In this field trial, the L3NM was used by the hierarchical controller (SDTN) to make the service requests to the IP Domain controllers (IP SDNc). For each service, the SDTN included network-specific information in the message sent to the domain controllers like transport LSP binding, routing profiles, or encryption profile. It assigned logical resources such as route targets or route distinguishers to keep the data synchronized between domains. \Cref{FIG:l3nm} shows the structure of the L3NM YANG data model, where the main container (VPN Service) is used to group the information of the VPN-Node (VRFs) and the VPN-Network Accesses (Interfaces).

@@

%\subsubsection{Network Topology}
%\label{subsection:IPtopo}

%The traditional SDN architecture relies on providing different levels of abstraction for each control layer. Therefore, the needs in terms of topology and knowledge of the service provider network differ among components.

%Network topology is an abstract representation of the physical nodes, links and network interconnections. It is crucial to get and graphically represent network information, such as:
%\begin{itemize}
%    \item Structure (Connectivity and Paths)
%    \item Performance (Available bandwidth per link)
%    \item Availability of physical and logical resources
%\end{itemize}

%Currently, the topology representations are limited to the scope of each of the network vendors i.e., each NMS has its particular/proprietary network view. Sometimes Dummy devices from a third party can be included to simulate the interconnection of the networks. However, nowadays obtain a unified view of the entire IP network is not possible.

%As part of the I2RS working group on the IETF defined a base model for an initial network topology representation has been defined. The model includes nodes and links. As a complementary work, several augmentations have been done over this base to cover the L2, L3, and TE functionalities.

%In this implementation, the topology models used to retrieve the IP network are: 
%\begin{itemize}
%\item IETF Network (RFC8345): Includes the Network, Node and Link concepts.
%\item IETF Network Topology (RFC8346): Includes Termination-Points inside the Nodes and IP specific parameters such as addressing information.
%\end{itemize}

@@

%i\uppercase{FUSION} is a reference model architecture defined to reinforce the network automation and programmability in a service provider environment. i\uppercase{FUSION} establishes a clear separation between network and service layers.  The i\uppercase{FUSION} main principles include the use of:

@@

\begin{itemize}
\item \textbf{SDN Domain}: It is a set of network elements under the supervision of the same SDN Controller. There are several possible levels in the decoupling of control and data planes. The preferred level of decoupling depends on the network technology. For example, network element runs the distributed protocols (e.g. IS-IS TE, RSVP-TE) in MPLS, and the controller only needs to configure it.

%\item \textbf{SDN Transport}: It is the whole network controlled by following SDN principles. It is divided into SDN Domains for technology/scalability/administrative principles. A SDN Transport Controller (also referred as SDN Orchestrator), will take care of stitching the different domains/layers/technologies.

@@

%\item \textbf{Southbound Interface}: It is the interface, based on a standard, between the SDN Domain Controller and the Network Element. Not only the communication protocol needs to be standard, but also the data model used.

%\item \textbf{Northbound Interface}: If is the interface, based on a standard, between the SDN Domain Controller and the OSSs and SDN Transport.
\end{itemize}

@@

%The SDTN entails the following functions: 
%\begin{enumerate}
%    \item End-to-end service binding and mapping.
%    \item End-to-end transport service abstraction.
%    \item End-to-end resource/topology discovery and composition.
%    \item End-to-end resource visualization.
%\end{enumerate}

%The SDN-Cs, on the other hand, are responsible of all the devices in the domain. Each SDN-C unifies the device configuration interface and provides vendor-agnostic network configuration, monitoring and resource discovery. Besides, the SDN-C exposes high-level network services abstraction to OSS and BSS layers through its NBI. Therefore, the abstraction of device-specific configuration from network service definition is one of the main features that the SDN-C implements. Moreover, the SDN-C entails the function of Path Computation Element (PCE) to manage and optimize traffic engineering (TE) in the domain.

@@

The goal SDN solution for the IP segment is composed of a single IP \replaced{SDNc}{SDN controller}, whose goal is to configure all the IP network elements. The target SBI for vendor-agnostic device configuration shall be compliant with NETCONF standard protocol. \added{OpenConfig started as a collaborative work between Google, some vendors and Service Providers, and now it is an industrial reference for device-specific configuration purposes. The philosophy behind OpenConfig is to add functionality to the device models based on the operator’s needs to avoid complex models which can not be implemented in real networks.} Due to its maturity level, the set of device configuration data models are the ones defined used in i\uppercase{FUSION} are OpenConfig.
%The IP SDN-C shall perform TE and PCE tasks. With that purpose, some standard and mature control protocols such PCEP and BGP-LS for MPLS networks, shall be implemented to complete the definition of the SBI. As a result, 
It is expected that the IP \replaced{SDNc}{SDN controller} will assume the control/management of:
\begin{itemize}
\item Device configuration of interfaces (VLANs) and routing protocols (BGP, ISIS, …)
\item Traffic Engineering of MPLS tunnels (LSPs). 
\item Overlay networks services (L2/L3 VPNs) device configuration (VRFs,\dots)
\end{itemize}

@@

%The NBI shall provide to higher entities within the SDN hierarchy:
%\begin{itemize}
%\item Device inventory information.
%\item A layered topology view (L2/L3, MPLS) of its controlled network entities.
%\item LSPs provisioning and PCE.
%\item Device abstraction for network services towards the SDTN, i.e., for overlay services VPNs (L2, L3)
%\item Network state and performance monitoring information of the IP domain. 
%\end{itemize}

@@

%The Hybrid SDN architecture proposed is compatible with a legacy control scenario where a distributed GMPLS control plane has already been deployed. GMPLS control plane can be centrally managed by a SDN domain controller by well-know and mature control protocols, such as PCEP, OSPF and/or BGP-LS already supported in GMPLS devices, benefiting the gradual introduction of SDN. However, current NMS solutions shall evolve to, or co-exist with, the SDN Controller model, enabling network programmability through its NBIs while keeping the current offered features for network creation, resources discovery and monitoring and service creation for L0/L1 layers. Standardization efforts targeting the definition of standard NBIs that can facilitate multi-vendor interoperability (by maintaining administrative domains for each vendor) such as ONF Transport API (T-API) \cite{lopez2016transport} and IETF models \cite{rfc8299} are the more promising definitions to implement such capabilities by abstracting the specific configuration of current distributed control planes embedded in Automatic Switched Optical Network (ASON) architectures. 
%i\uppercase{FUSION} relays on ONF Transport API 2.1 as as the reference NBI for the SDN implementation in the optical transport segment, having been experimented in several proof of concepts \cite{mayoral2016first}.

In the medium and long term, the direct programmability of the components can have interest in Point-To-Point, Metro and Regional scenarios, where disaggregation of optical transceivers and line side components can play an important role. In this line, OpenROADM \cite{oda2016learning} and OpenConfig projects have already defined device configuration models for transponders and open line systems. 
%Telefonica is approaching this transformation of the optical control in two phases:

%\begin{enumerate}
%    \item \textbf{Partial disaggregation}: As a medium term objective, where the target is to define a standard interface based on NETCONF/YANG, which allows of the Optical SDN-C to manage third-party terminal devices (i.e., transponders) that can transmit over the vendor line system.

%    \item \textbf{Full disaggregation}: The long term objective is the open management of the line system, i.e., the defragmentation of the optical transport network in vendor islands by the adoption of a common standardized interface for open line systems (multi-vendor) to be managed by a single optical SDN-C.
%\end{enumerate}
%In this section the reference standard model use for the control/management of networks based on WDM/OTN technologies is described. 
%The proposed solution for the Optical SDN-C's NBI is based on ONF T-API information models \cite{mayoral2016first}.  NBI message transportation is performed using the RESTCONF protocol. Please note, that in this transport domain, the same solution (T-API + RESTCONF) is also proposed to be the NBI of the hierarchical SDTN Controller towards Service B/OSS layers.

%\subsubsection{Provisioning}
%
%\begin{figure*}
%	\centering
%		\includegraphics[scale=0.75]{figs/ONF-T-API.png}
%	\caption{Multi-layer topology and connectivity models based on ONF T-API}
%	\label{FIG:ONF-T-API}
%\end{figure*}

%The connectivity provisioning into the WDM/OTN optical layers is strictly related with the model proposed. \ref{FIG:ONF-T-API} represents the optical network topology arrangement proposed in the previous figure and the multi-layer hierarchical connectivity modelling which is being detailed in this section.

%The connectivity model introduces to central concepts. On one hand, the tapi-connectivity:connectivity-service  object models "intended" solicited by the T-API client layer, to be deployed into the network. On the other hand, the tapi-connectivity:connection objects represents the actual adjacencies between logical interfaces created by configuring the network, in brief the connectivity configurations in the network.

%The network model is multi-layer so it is needed to establish multi-layer relationships into the model. These relationships are construct by the relationship between Logical Termination Points (LTPs) which in T-API are modelled as two different objects: the Node-Edge-Points (NEPs) which represents the resources available at a give layer LTP to provide connections, and the \textbf{Connection-End-Points (CEPs)} which consumes part of all of the NEPs exposed resources to create the connections between different points of the network. In our proposed reference implementation, we distinguish between two different level of connections:

%\begin{itemize}
%    \item \textbf{Cross-Connections (XC)}: defined as a connection between Connection-End-Points of the same layer within a Forwarding-Domain (represented as a\\ \texttt{tapi-topology:node} object). 
%    \item \textbf{Top Connections}: are defined as end-to-end connections between CEPs within the same layer which may span multiple Forwarding-Domains. Top connections are composed by zero or more XCs which belong to the same layer of the Top Connection.
%\end{itemize}

%Then, the multi-layer relationships are construct by stacking CEPs over NEPs and providing the upper layer resources representations by the dynamic creation of new NEPs, e.g., an OTSi connection, when created and operational, it provides the ODU upper layer resources in the form of NEPs which in turn can be consumed to create ODU connections which will provide the DSR layer resources.

%\subsubsection{Topology}
%\label{subsection:OPTopo}

%The topology model should provide the explicit multi-layer topology representation of the L2-L0 network including OTS, OMS, MC, OTSIMC, OTSi/OTSiA, ODU, DSR layers. The network logical abstraction collapses all network layers (DSR, ODU, OTSi/OTSiA and Photonic Media (OTSiMC, MC, OMS, OTS)) which are represented explicitly into a single topology (T0 – Multi-layer topology), modelled as a \texttt{tapi-topology:topology} object within the: \\
%\texttt{tapi-topology:topology-context/... \\
%... tapi-topology:nw-topology-service} and \\ \texttt{tapi-topology:topology-context/topology}.

%The T0 – Multi-layer topology MUST include:

%\paragraph{DSR/ODU Layers:}
%DSR/ODU forwarding domains represented as multi-layer and multi-rate \texttt{tapi-topology:node}, allowing the representation of the internal mapping between DSR and ODU NEPs (multi-layer) and the multiplexing/de-multiplexing across different ODU rates (multi-rate).

%The DSR/ODU layer network MUST be represented explicitly at the lowest partitioning level, i.e., each DSR/ODU forwarding domain MUST be represented as a single tapi-topology:node. The following network components included within the category of ODU forwarding domain are:

%\begin{itemize}
%    \item Transponders.
%    \item Muxponders.
%    \item OTN switching nodes connecting client and line %boards.
%\end{itemize}

%\paragraph{OTSi/Photonic Media layers:}
%The OTSi layer represents the optical side of the optical terminals (transponders/muxponders). This layer consists of nodes representing the mapping and multiplexing of OTSi signals. It consists of nodes including OTSi client endpoints representing the Trail Termination Points (TTPs) of the OTSi connections and OTSi/OMS endpoints representing the physical connectivity with ROADM/FOADM add/drop ports.

%DSR/ODU and OTSi layers may be collapsed into a single multi-layer node or split into two logical nodes representations by using the Transitional links concept to represent the potential layer transitions between ODU and OTSi.

%\paragraph{Photonic-Media layer:}
%The Photonic-Media layer models the Optical Line Protection  (OLP) components, the Reconfigurable/Fixed Optical Add Drop Multiplexers (ROADM/FOADM) and In-Line Amplifiers (ILAs) network elements. Moreover, all the lowest photonic connectivity is represented as PHOTONIC\_MEDIA \texttt{tapi-topology:link} objects collapsing OTS/OMS layers and allowing placing specific monitoring OAM functions of these layers. These forwarding domains SHALL expose the capability of create Media Channel connection and connectivity services between its endpoints.

%\begin{figure}
%	\centering
%		\includegraphics[scale=0.75]{figs/Media_channel.png}
%	\caption{Media-channel entities relationship}
%	\label{FIG:Media_channel}
%\end{figure}

%Moreover, the Media Channel layer the available resources for the reservation of  spectrum resources for a given OTSi channel. The concatenated reserved portion of the along a route is represented as a Media-Channel (MC) construct, and by the OTSiMC construct which represents the actual portion of the spectrum occupied by the signal (MC spectrum must be wider than the OTSiMC). These modelling concepts are critical for the realization of the Open Line System concept introduced by the partially disaggregation of the optical networks. See \ref{FIG:Media_channel} graphical representation for more clarity.

\subsection{Integration of SDTN \added{Controller} in the overall operator’s systems architecture}
\label{section:sdtn}
One of the main reasons for deploying an SDTN Controller is service automation. It facilitates that manual services and network configurations become automated and available through its NBIs, enabling the network automation progressively. As the main design principle, the abstraction level provided by the SDTN \added{Controller} can be different based on the needs of consumers. Thus, the information exported through the NBI towards OSS and other platforms will cover several functional areas with several levels of abstraction: (1) network topology, (2) service provisioning, (3) performance management, (4) network planning and design, and (5) fault management. 
%\begin{itemize}
%    \item \textbf{Network topology}: An hierarchical topology view of the network % resources, based on the OSI layers.
%    \item \textbf{Service’s provisioning}: The available set of network services % to different clients.
%    \item \textbf{Performance Management (PM)}.
%    \item \textbf{Network Planning and Design (NPD)}.
%    \item \textbf{Fault Management (FM)}.
%\end{itemize}

% Please add the following required packages to your document preamble:
% \usepackage[table,xcdraw]{xcolor}
% If you use beamer only pass "xcolor=table" option, i.e. \documentclass[xcolor=table]{beamer}

%\begin{table}[]
%\caption{Set of use cases Tested in the POC}
%\label{TAB:use_cases}
%\begin{adjustbox}{width=0.5\textwidth}
%\begin{tabular}{|
%>{\columncolor[HTML]{EFEFEF}}l |l|c|}
%\hline
%\cellcolor[HTML]{C0C0C0}\textbf{Fuctional Requeriment} & \cellcolor[HTML]{C0C0C0}\textbf{Type}      & \multicolumn{1}{l|}{\cellcolor[HTML]{C0C0C0}\textbf{Priority}} \\ \hline
%Network topology collection                            & L2 IP Topology Collection                  & \cellcolor[HTML]{ECF4FF}Must                                   \\ \hline
%Network topology collection                            & L3 IP Topology Collection                  & \cellcolor[HTML]{ECF4FF}Must                                   \\ \hline
%Service’s provisioning                                 & L3VPN Creation / Deletion / Modification   & \cellcolor[HTML]{ECF4FF}Must                                   \\ \hline
%Service’s provisioning                                 & L2VPN Creation / Deletion / Modification   & \cellcolor[HTML]{ECF4FF}Must                                   \\ \hline
%Performance Management (PM)                            & Collect counters from the network          & Planned                                                        \\ \hline
%Network Planning and Design (NPD)                      & Collect counters from the network          & Planned                                                        \\ %\hline
%Fault Management (FM)                                  & Collect alarms and events from the network & Planned                                                      \\ %\hline
%\end{tabular}
%\end{adjustbox}
%\end{table}

@@

%On its SBI, the SDTN will integrate with the domain controllers; Each Transport SDN-C shall expose vendor-agnostic network-level programmability and resource discovery functionalities. So the SDTN will be able to perform the correct data integration and exposure.

%On the SBI of the SDTN, each technology Transport SDN-C shall expose vendor-agnostic network level programmability and resource discovery functionalities. The SDTN's SBI is intended, but not limited, to provide access to device’s configuration data, to expose per-OSI layer topology and network inventory information, and to offer active monitoring of device configuration changes and network state data (i.e., traffic statistics). Alarm and device inventory information for FM and RIM respectively, is intended to be managed at the SDN-C in a first phase, but its exposure through the SDTN will be evaluated too.

%\section{Network Programmability}
%\label{section:net}

%The network management has lagged behind other technologies quite drastically. In more than 20 years, there has not been radical improvements in the device management area. One of the major advances was related to the usage of SSH instead of TELNET for secure implementations. However, scripts based on EXPECT are the most programmatic way to access the devices in the service providers across the globe.

%This lack of manageability is often better understood when networking is compared with other technologies. For example, hypervisor managers, wireless controllers, IP PBXs, PowerShell, DevOps tools are part of the continous Integration and continous Development (CI/CD) in cloud providers \cite{mittal2017cloud,demchenko2016zerotouch}. Some of these are tightly coupled from a certain vendor, but others are more loosely aligned to allow for multi-platform management, operations, and agility \cite{edelman2018network}.

%To allow the network programmability it is completely necessary to use SDN as the reference for the introduction of network APIs. It must be taken into account that these network APIs are not just related to the service provision automation (provision is a repetitive task and its components can be described in a technological way). The usage of APIs and programmatic interfaces can automate and offer much more than pushing configuration parameters.

%It can be used to cover all the FCAPs (Fault, Configuration, Accounting and Performance) capacities and Network planning tasks:
%\begin{itemize}
%    \item Network planning tasks if we include APIs to export distributed information such as the RIB, FIB, and TE-Databases.
%    \item Used to deploy close-loop decision systems to take actions based on events reported by the devices.
%    \item Automatically visualize the network relationships between the IP/MPLS and Optical packet transport domain. Creating a common network view. 
%\end{itemize}

%\section{Service Models}
%\label{section:models}

%Due to the network programmability necessities. Protocols and Data models must be selected to allows the interaction between the network controllers and devices. There has been a great interest in using YANG to define these data models. However, this YANG-based set of definitions may include two groups of models. One set used strictly to attack the devices and the ones used to describe services in a portable and technological way (independent of which network operator uses the model). 
%This differentiation between Service and Device Models introduced in early 2018, would allow abstract some information in the control layer. The service models may be used as part of the SDN architecture, describing the data interchanged between the network controllers and between OSS systems and the control Layer \cite{wu2017service}. While the device models described the specific configuration of the network elements.

%Example of service models definitions used during this implementation are described in \ref{section:IPmodels} for IP and \ref{section:OPTmodels} for Optical.

@@

%\begin{figure}
%	\centering
%		\includegraphics[width=\linewidth]{figs/multidomain_service_provisioning%_workflow.png}
%	\caption{Workflow for multi-domain service provisioning SDTN-SDN-C}
%	\label{FIG:multidomain_service_provisioning_workflow}
%\end{figure}

%\subsubsection{Multi-layer Topology}

%The multi-layer topology use case is based on the data provided by all the IP and Optical WDM/OTN SDN-Cs. The scope of it, includes the composition of multiple sources and data-formats (i.e IETF context for IP or T-API context for Optical) to create a common view of the network. The models used to create the multi-layer topology were:
%\begin{itemize}
%    \item For the optical domain representation: Topology and Connectivity Service modules. Those models provides information for layers from L0 to L2.
%    \item For the IP/MPLS domain representation: The IETF\\ \texttt{ietf-network:networks} is the basis to expose the network model(nodes and IP Links), and the \\ \texttt{.../node/nt:termination-point}
%    were used to expose the Terminations-points (Ports) of a specific node.
%\end{itemize}

%Additionally, a correlation parameter is defined to join the IP and Optical layers. We have denoted this parameter as the \texttt{Plug-id}. This additional definition is needed due to there is no a dynamic protocol or common element between the ONF/IETF standards that would allow their direct correlation. The \texttt{Plug-id} parameters is added in the IETF termination points and in the T-API connectivity services.

%It is necessary to remark in the T-API model side the \texttt{Plug-id} parameter is a "String" and in the IETF side is "Binary", therefore the SDTN MUST have the ability to traduce the “Binary” value to “String” or vice versa, to be able to make the match in the correlation process considered to be standardized.

%To fill the \texttt{Plug-id} attribute automatically between the layers, it is required that the SDTN performs a process based on meta-heuristic algorithms. In which, the performance data at termination-points level both in the IP and in the Optical layers will be correlated. Clock \& Time must be perfectly synchronized between the Tx/Rx points, bidirectionally. Meta-heuristics must include the geo-location parameter of the nodes as input. However, the design of this process is currently under study and is based on real-time information that can be delivered by the telemetry engines.

%\begin{figure}
%	\centering
%		\includegraphics[width=\linewidth]{figs/ip_topology_workflow.png}
%	\caption{Message Interchange for IP Topology Discovery between the SDTN and IP SDN-C}
%	\label{FIG:ip_topology_workflow}
%\end{figure}

%Starting with the Optical Domain, a set of T-API version 2.1 queries were sent in order to build the topology by extracting the list of networks as well as topology details such as nodes, links, connections and service-interface-points available. The query exchange process between is shown in the UML diagram depicted in \ref{FIG:optical_topology_workflow}.

%\begin{figure}
%	\centering
%		\includegraphics[width=\linewidth]{figs/optical_topology_workflow_2.png}
%	\caption{Messages Interchanged for Optical Topology Discovery between the SDTN and Optical SDN-C}
%	\label{FIG:optical_topology_workflow}
%\end{figure}

%Regarding the IP Domain, to obtain the whole network topology, a query to the controller’s NBI to retrieve the network topologies available for all layers is sent; from here on, the client must perform per-layer queries to get more detailed information such as nodes, termination points of nodes, links, etc. \ref{FIG:ip_topology_workflow} displays the RESTCONF based queries in UML format for the topology retrieval in the IP Domain. 
%At this point, no inter-domain links were retrieved since there is no automatic mechanism supported on the domain controllers to expose information such as TTI (on the Optical Domain), LLDP (on the IP Domain) or \texttt{Plug-id} (for both domains); therefore \texttt{Plug-id} based link discovery was simulated by adding user defined \texttt{Plug-id} values to each port or at least to each domain edge port manually with the use of Python scripting. The inputted plug-id data was automatically detected by the SDTN and the inter-domain links between ports with matching \texttt{Plug-ids} were created, resulting in a complete multi-layer/multi-domain end-to-end topology as seen from the SDTN GUI.

%\begin{figure}
%	\centering
%		\includegraphics[width=\linewidth]{figs/topology_workflow.png}
%	\caption{Workflow for multi-layer topology SDTN-SDN-C}
%	\label{FIG:topology_workflow}
%\end{figure}

%\begin{figure*}
%	\centering
%		\includegraphics[scale=1]{figs/field_trial_environment_ip.pdf}
%	\caption{Network Plane of the Field Trial Environment for IP/MPLS Testing and Evaluation}
%	\label{FIG:field_trial_ip}
%\end{figure*}

%\begin{figure*}
%	\centering
%		\includegraphics[scale=1]{figs/field_trial_environment_optical.pdf}
%	\caption{Network Plane of the Field Trial Environment for Optical/WDM Testing and Evaluation}
%	\label{FIG:field_trial_optical}
%\end{figure*}

@@

In this work, we have deployed the i\uppercase{Fusion} architecture in a Telefonica Colombia field trial environment. The field trial includes not just the i\uppercase{FUSION} SDN control layer (\Cref{sec:contollay}), but also the network elements described in \Cref{sec:netlay}.%, which is a multivendor IP/MPLS network using an underlying optical infrastructure.

@@

%In other hand, from the Optical perspective, SDN-Cs follow a similar integration. At SBI, OpenROADM + OpenConfig models are used on top of NETconf/YANG protocol. At NBI, a T-API v2.1 implementation is used.

@@

Thus, the HL3 routers from each region establish an eBGP session against the Core-Routers (HL2). This session exports the Router-ID plus Label information of all the routers in the region using BGP label unicast (LU). Additionally, another eBGP session between the HL3 of the region and the core Router-Reflectors to export the VPNv4 routes from each VPN service. This eBGP session requires a mandatory Next-Hop-Unchanged configuration to avoid network loops or misconfigured paths. %All of this control plane setup allows the creation of an end-to-end LSP from the access layer to the platforms without changing the configuration during the service provisioning.

%Additionally, to deploy any of this service the network has to fulfill the following basic requirements established between origin and destination:
%\begin{itemize}
%    \item PEs connectivity based on IGP Router-ID/Loopback reachability.
%    \item Label switching protocol enabled. MPLS and Labelling mechanism LDP, RVSP, other. 
%    \item MP-BGP sessions between the PEs (address-family vpnv4/6, ipv4/6).
%    \item Virtual Routing network instance. 
%\end{itemize}

Two-vendor WDM infrastructures transported the IP/MPLS links of vendor B HL2s and vendor A HL3s. The optical transport infrastructure was built as two independent metropolitan optical networks. We used (4) four nodes in each optical network with 100Gb and 10Gb lambdas for these experiments. %\ref{FIG:field_trial_ip} illustrates the optical WDM part of the field trial enviroment.

%\ref{TAB:tested_use_cases} shows the uses cases tested on the network scenario, these have been classified into three main categories: Topology discovery, IP service provisioning and Optical service provisioning. A general overview and the results obtained for each of them are approached individually in further sub-sections of this chapter. Regarding the single-domain use cases, results are shown independently for each vendor and for each network domain in particular.

%\begin{table*}
%	\caption{List of Multi-Layer Multi-Domain Tested Use Cases}
%	\centering
%		\includegraphics[scale=0.5]{figs/tested_use_cases.png}
%	\label{TAB:tested_use_cases}
%\end{table*}

%\subsection{Multi-layer Topology Discovery \& Visualization}
%Obtaining the end-to-end multi-layer and multi-domain topology via automatic network discovery is considered the starting point towards the deployment of the SDTN multi-layer solution in the network scenario described in Section 5. The first and most important step is for SDTN to discover all the network elements of the different domains by using the APIs provided by each of the IP and Optical Controllers, as described in previous sections.

%Technically talking, the following steps take part of the end-to-end network discovery procedure:
%\begin{enumerate}
%    \item \textit{Enabling Network Adapters}: The network for each domain was discovered from the domain controllers individually by using Domain Adapters; each adapter uses the API provided by the Domain Controller to retrieve data such as nodes, inventories, termination points, links and services within it.
%    \item \textit{Data Model Mapping}: The information model used in the Domain Controller's API is mapped into the SDTN model in order to harmonize the data across all domains, providing both a per-layer view within the inter-layer links and client-server relationships, thus resulting in a complete multi-layer view of the network and its services.
%    \item \textit{Inter-domain links discovery}: End-to-end view of the whole network topology is formed by discovering the inter-domain links. Those interconnect the different vendors into a whole end-to-end multi-layer and multi-domain topology. There are many different mechanisms\footnote{TTI or plug-id for L0/L1, LLDP for L2, IP address based discovery for L3, among others such as alarm/statistics correlation with or without test data/circuits, inter-domain links simulation if no automatic mechanism is available} 
%    to achieve this purpose, such as TTI for OTU links, IP addresses for IP links etc. If all the data required for full inter-domain and inter-layer link discovery is not reported by the third-party controllers, external data can be fed in via the SDTN NBI. 
%\end{enumerate}

>>>>

\author{
\IEEEauthorblockN{Samier Barguil\IEEEauthorrefmark{1}, Victor Lopez\IEEEauthorrefmark{2}, Cristyan Manta-Caro\IEEEauthorrefmark{4}, Arturo Mayoral Lopez De Lerma\IEEEauthorrefmark{2},\\
Oscar Gonzalez De Dios\IEEEauthorrefmark{2}, Edward Echeverry\IEEEauthorrefmark{3}, Juan Pedro Fernandez-Palacios\IEEEauthorrefmark{2}, Janne Karvonen\IEEEauthorrefmark{5},\\
Jutta Kemppainen\IEEEauthorrefmark{5}, Natalia Maya\IEEEauthorrefmark{5}, and Ricard Vilalta\IEEEauthorrefmark{6}}
\IEEEauthorblockA{
\IEEEauthorrefmark{1}Universidad Autonoma de Madrid, Madrid, Spain\\
\IEEEauthorrefmark{2}Telefonica I+D, Ronda de la Comunicacion, Madrid, Spain \\
\IEEEauthorrefmark{3}Telefonica Movistar, Transversal 60 No 114ª -55. Bogotá, Colombia\\
\IEEEauthorrefmark{4}Wipro Technologies Ltd., Doddakannelli, Sarjapur Road
Bengaluru - 560 035, India\\
\IEEEauthorrefmark{5}Infinera Corporation, 140 Caspian Court, Sunnyvale, CA 94089, USA\\
\IEEEauthorrefmark{6}Centre Tecnol\`{o}gic de Telecomunicacions de Catalunya (CTTC/CERCA), Castelldefels, Spain}}

@@

\begin{abstract}
Network operators can not change their footprint to upgrade the entire Network to support SDN from scratch. \replaced{Hence,}{This why} network operators adapt\added{ed} the original SDN concept into a hybrid SDN approach to have a pragmatic, evolutionary, and economically viable solution. This paper tests an SDN architecture based on a hierarchical structure of a Software-Defined Transport Networking (SDTN) controller, which deals with the end-to-end service provision and topology collection on top of a set of SDN controllers for each technological domain. The authors demonstrate L3/L2 VPN service deployment using an underlay multi-domain IP/Optical networks in a field trial. The approach allows a service provider to migrate brownfield scenarios into SDN domains using programmatic interfaces.

\end{abstract}

@@

@@

@@

@@

\begin{itemize}

\item \textbf{SDN Domain}: It is a set of network elements under the supervision of the same SDN Controller. There are several possible levels in the decoupling of control and data planes. The preferred level of decoupling depends on the network technology. For example, network element runs the distributed protocols (e.g. IS-IS TE, RSVP-TE) in MPLS, and the controller only needs to configure it.

@@

\end{itemize}

@@

@@

The goal SDN solution for the IP segment is composed of a single IP \replaced{SDNc}{SDN controller}, whose goal is to configure all the IP network elements. The target SBI for vendor-agnostic device configuration shall be compliant with NETCONF standard protocol. \added{OpenConfig started as a collaborative work between Google, some vendors and Service Providers, and now it is an industrial reference for device-specific configuration purposes. The philosophy behind OpenConfig is to add functionality to the device models based on the operator’s needs to avoid complex models which can not be implemented in real networks.} Due to its maturity level, the set of device configuration data models are the ones defined used in i\uppercase{FUSION} are OpenConfig.

It is expected that the IP \replaced{SDNc}{SDN controller} will assume the control/management of:
\begin{itemize}
\item Device configuration of interfaces (VLANs) and routing protocols (BGP, ISIS, …)
\item Traffic Engineering of MPLS tunnels (LSPs). 
\item Overlay networks services (L2/L3 VPNs) device configuration (VRFs,\dots)
\end{itemize}

@@

@@

In the medium and long term, the direct programmability of the components can have interest in Point-To-Point, Metro and Regional scenarios, where disaggregation of optical transceivers and line side components can play an important role. In this line, OpenROADM \cite{oda2016learning} and OpenConfig projects have already defined device configuration models for transponders and open line systems.

\subsection{Integration of SDTN \added{Controller} in the overall operator’s systems architecture}
\label{section:sdtn}
One of the main reasons for deploying an SDTN Controller is service automation. It facilitates that manual services and network configurations become automated and available through its NBIs, enabling the network automation progressively. As the main design principle, the abstraction level provided by the SDTN \added{Controller} can be different based on the needs of consumers. Thus, the information exported through the NBI towards OSS and other platforms will cover several functional areas with several levels of abstraction: (1) network topology, (2) service provisioning, (3) performance management, (4) network planning and design, and (5) fault management.

@@

@@

@@

In this work, we have deployed the i\uppercase{Fusion} architecture in a Telefonica Colombia field trial environment. The field trial includes not just the i\uppercase{FUSION} SDN control layer (\Cref{sec:contollay}), but also the network elements described in \Cref{sec:netlay}.

@@

@@

Thus, the HL3 routers from each region establish an eBGP session against the Core-Routers (HL2). This session exports the Router-ID plus Label information of all the routers in the region using BGP label unicast (LU). Additionally, another eBGP session between the HL3 of the region and the core Router-Reflectors to export the VPNv4 routes from each VPN service. This eBGP session requires a mandatory Next-Hop-Unchanged configuration to avoid network loops or misconfigured paths.

Two-vendor WDM infrastructures transported the IP/MPLS links of vendor B HL2s and vendor A HL3s. The optical transport infrastructure was built as two independent metropolitan optical networks. We used (4) four nodes in each optical network with 100Gb and 10Gb lambdas for these experiments.
